---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- You need to render `README.Rmd` to keep `README.md` up-to-date. -->

<!-- use`devtools::build_readme()` for this.  -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kitems

<!-- badges: start -->

<!-- badges: end -->

The goal of kitems is to provide a framework to manage data.frame like
items and a set of tools to implement it within a shiny web application.
For that reason, it is delivered as a shiny module.

## Code coverage

[![codecov](https://codecov.io/gh/thekangaroofactory/kitems/graph/badge.svg?token=7P74NK51JJ)](https://codecov.io/gh/thekangaroofactory/kitems)

## Installation

You can install the development version of kitems from
[GitHub](https://github.com/thekangaroofactory/kitems) with:

``` r
# install.packages("devtools")
devtools::install_github("thekangaroofactory/kitems")
```

## Demo

A basic shiny demo app which shows you how to use the package is
delivered and can be run:

```{r demo, eval=FALSE, include=FALSE}
library(kitems)

runExample()

```

## Framework Specifications

The kitems framework is based on a two main notions: data model and
items.

### Data Model

The data model contains the specifications of the items you want to
manage. It includes:

-   attribute names,
-   attribute types,
-   default values,
-   default functions,
-   filter,
-   skip.

Attribute names and types are obvious. Supported data types are:

-   numeric,
-   integer,
-   logical,
-   character,
-   factor,
-   Date,
-   POSIXct,
-   POSIXlt.

### Items

Items is a data.frame containing the data to be managed.

## Reactive Values

kitems strongly relies on Shiny (shiny and shinydashboard packages are
set as 'depends' dependencies). It provides reactive values (accessible
from the r object passed as an argument of the module server):

-   r[[data_model]]
-   r[[items]]
-   r[[filtered_items]]
-   r[[selected_items]]
-   r[[filter_date]]
-   r[[trigger_add]]
-   r[[trigger_update]]
-   r[[trigger_delete]]
-   r[[trigger_save]]

> [!CAUTION] Data model and items reactive values should be handled with
> caution as updating there values will trigger auto save (for items, if
> autosave = TRUE).

### Data Model

```{r data_model, eval=FALSE}
library(kitems)

# -- define module id
id <- "mydata"

# -- get data model reactiveValue name
data_model <- dm_name(id)

```

The data model for this id can be accessed **in a reactive context** by
using: r[[data_model]]()

### Items

```{r items_name, eval=FALSE}
# -- define module id
id <- "mydata"

# -- get data model reactiveValue name
items <- items_name(id)

```

The items for this id can be accessed **in a reactive context** by
using: r[[items]]()

### Triggers

Triggers are objects that can be used to communicate actions to the
module server:

-   trigger_add: add an item to the item list
-   trigger_update: update an item from the item list
-   trigger_delete: delete an item from the list
-   trigger_save: save the item list (if autosave is turned off)

## Data model

### Initialization

Several options are available to initialize a data model when starting
the module server:

-   pass a data model to data.model argument,
-   keep the data.model argument NULL.

If data.model is set, it will be used to initialize the persistent data
model. Note that once initialized, the persistent data model can be
updated (for example through the admin UI) and will have the priority
over the data.model argument.

If data.model is NULL, the admin UI will display a create button to
start a new data model, as well as an import data button.

### Checking integrity

When the module server is launched, it will perform an integrity check
to ensure that the data (items) and data model are aligned.

This is done through the dm_check_integrity() function. If not aligned,
the data model will be updated to match with any missing attribute for
example.

## Item management

### Item creation

It's recommended to use item_create() to create the item to be added to
the item list:

```{r item_create, eval=FALSE}
library(kitems)

# -- define module id
id <- "mydata"

# -- get names
data_model <- dm_name(id)
trigger_add <- trigger_add_name(id)

# -- define inputs and create item (waning! this should be used in a reactive context)
input_values <- data.frame(id = 1, text = "demo")
item <- item_create(values = input_values, data.model = r[[data_model]]())

# -- trigger action (waning! this should be used in a reactive context)
r[[trigger_add]](item)
```

Note: if autosave has been turned off, r[[trigger_save]]() should be
used to make item changes persistent.

### Item update

It's recommended to use item_create() to create a new item and replace
the one in the item list:

```{r item_update, eval=FALSE}
library(kitems)

# -- define module id
id <- "mydata"

# -- get names
data_model <- dm_name(id)
trigger_update <- trigger_update_name(id)

# -- define inputs and create item (waning! this should be used in a reactive context)
input_values <- data.frame(id = 1, text = "demo")
item <- item_create(values = input_values, data.model = r[[data_model]]())

# -- trigger action (waning! this should be used in a reactive context)
r[[trigger_update]](item)
```

### Item delete

To delete an item, just pass it's id to the trigger:

```{r item_delete, eval=FALSE}
library(kitems)

# -- define module id
id <- "mydata"

# -- get trigger name
trigger_delete <- trigger_delete_name(id)

# -- trigger action (waning! this should be used in a reactive context)
item_id <- 1704961867683 
r[[trigger_delete]](item_id)

```

## Views

### Default view

The default view is provided through items_view_DT(). It is based on the
r[[items]] content (with data model masks applied).

### Filtered view

The filtered view is provided through items_filtered_view_DT() It is
based on the r[[filtered_items]] content (with data model masks
applied).

## Selected Item(s)

Both default and filtered view have row selection enabled. Selecting
row(s) in *one or the other* table will trigger the update of
r[[selected_items]] reactive value.

> [!CAUTION] It is not recommended to display both views in the same UI
> section as only the last selection will be kept in r[[selected_items]]

Selected item(s) will also trigger which buttons are available (or
NULL).

## Inputs

### Date sliderInput

If the data model has an attribute named 'date', a date sliderInput will
be created automatically to enable date filtering.

This sliderInput will trigger a filter on the items to be displayed in
the filtered view, i.e. only the matching items will be part of this
view.

If not implemented through date_slider_INPUT() in your application's UI,
then no date filter is applied by default on the filtered view.

Filter value (the sliderInput range) is available in the
r[[filter_date]]() reactive value.

**Note** In case you want to build a reactive object on top of
r[[filter_date]] and r[[filtered_items]], it is recommended to test both
for NULL to avoid firing computation at startup.

### Buttons

-   create_BTN: create item
-   update_BTN: update selected item (single selection)
-   delete_BTN: delete selected item(s) (supports multi-selection)

update_BTN and delete_BTN return NULL if no row is selected.

## UI objects

### Admin UI

In order to avoid code duplication, an admin UI interface is provided by
the admin_ui() function.

```{r admin_ui, eval=FALSE}

# -- Inside tabItems
tabItems(...,

         tabItem(tabName = "mydata",
                 
                 # -- Admin UI
                 fluidRow(
                   column(width = 12,
                          
                          kitems::admin_ui("mydata"))))
)

```

### Sidebar

A sidebarMenu object (menuItem) is also available as an output of the
dynamic_sidebar() function. It will contain one menuSubItem object per
data model (stored in the r reactive values).

To implement it in your sidebar UI, it must first be rendered in your
application server code:

```{r sidebar_server, eval=FALSE}
# -------------------------------------
# Generate dynamic sidebar
# -------------------------------------

output$menu <- renderMenu(dynamic_sidebar(r))

```

Then you can use this output inside the application UI code:

```{r sidebar_ui, eval=FALSE}
# -------------------------------------
# Define Sidebar UI
# -------------------------------------

sidebar <- dashboardSidebar(

  # -- static section
  sidebarMenu(
    menuItem("Home", tabName = "home", icon = icon("dashboard"), selected = TRUE)),

  # -- add dynamic section
  sidebarMenu(id = "tabs", sidebarMenuOutput("menu")),

  collapsed = FALSE)

```

The tabName linked to each menuSubItem should be named after the module
id. Here is an example how to access the administration console from the
menuSubItem:

```{r tabname_ui, eval=FALSE}
# -- kitems admin
tabItem(tabName = "my_data",
        
        # -- Admin UI
        fluidRow(
          column(width = 12,
                 
                 kitems::admin_ui("my_data"))))
```

## Import & migration

When the module server is initialized with data.model = NULL, an import
data button is displayed in the data model section of the admin UI. This
allows to create a data model from existing data.

## Reactivity

### Observe items

```{r observer_items, eval=FALSE}
# -------------------------------------
# Observe item lists
# -------------------------------------

id <- "mydata"
items <- items_name(id)

# -- Module id = mydata
observeEvent(r[[items]](), {
  
  cat("Main application server observer: items object has just been updated. \n")
  
})

```

Note: as long as you know the module id, you can shortcut it with
observeEvent(r\$mydata_items(), ...) but this makes your code less
resilient to changes.
