---
title: "Environment"
format: html
editor: visual

vignette: >
  %\VignetteIndexEntry{Environment}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

This article is here to keep all environment considerations in a single place.

## Path

Both data model & item files are stored in a specific folder dedicated to a given instance of the module server.

-   The folder is named after the `id` of the module

-   It is located under the path given to the `path` argument

```{r, eval=FALSE}
kitems::kitems(id = "mydata", path = "./data")
```

For this reason, unless there is a single instance of the module, it is not recommended to name any of the `id` with same name as the last folder of the provided path. It may result in files being stored at different levels.

Examples:

```{r, eval=FALSE}
kitems::kitems(id = "data", path = "./path/to/data")
```

Files will be stored in “./path/to/data”

```{r, eval=FALSE}
kitems::kitems(id = "data_1", path = "./path/to/data")
```

Files will be stored in “./path/to/data/data_1”

## Debug

For debugging purpose, the package implements traces through outputs to the console.

But because printing to the console has serious impact on the overall app performance, it has been decided to wrap the trace mechanism into a function that only outputs the log when it is activated.

To set the trace level, you need to use the `ktools::trace_level()` function.
